---
title: "ggplot vs seaborn - R"
author: "JB"
date: "8 December 2015"
output: html_document
---

This is the R version of the code for examining the meat dataset.

```{r}
meat = read.csv('/Users/Jake/Projects/ggplotvsseaborn/meat.csv')
str(meat)
```

Coerce date to an actual date object:

```{r}
meat$date = as.Date(meat$date)
```

#Data

```{r, echo=FALSE}
head(meat, n=10)
tail(meat, n=10)


```

This is the sum by year aggregation:

```{r}
library(lubridate)
library(dplyr)
#First, create a year column using the lubridate package:
meat = mutate(meat, year = year(date))
#Then, group by year and sum columns:

meat_by_year = meat %>%
  group_by(year) %>%
  summarise_each(funs(sum))

#clean to get just the columns we are interested in:

meat_by_year = select(meat_by_year,
                      year, beef, veal, pork, lamb_and_mutton, broilers, other_chicken, turkey)
```

#Plotting:

```{r}
#Beef yearly trends:
library(ggplot2)
ggplot(meat_by_year, aes(x = year, y = beef)) + 
  geom_bar(stat = 'identity') +
  scale_y_continuous() +
  ggtitle('Head of Cattle Slaughtered by Year')
        
```

#Melt the data so we can see trends across different types of livestock:

```{r}
library(reshape2)

#we focus on beef, veal, pork and turkey.
meat_by_year = select(meat_by_year,
                      year, beef, veal, pork, turkey)

melt_data = melt(as.data.frame(meat_by_year), id.vars = 'year')

ggplot(melt_data, aes(x= year, y = value, color = variable)) + 
  geom_line() +
  ggtitle('By Year, by type of Meat')

#alternatively, using faceting

ggplot(melt_data, aes(x = year, y = value)) + geom_line() +
  facet_wrap(~ variable)


#Change the colurs up a bit - wesanderson pallettes:
library(wesanderson)
ggplot(melt_data, aes(x= year, y = value, color = variable)) + 
  geom_line() +
  scale_fill_manual(values = wes_palette("Darjeeling2")) +
  ggtitle('By Year, by type of Meat')

```

#Trends over time, not aggregated

```{r}
meat_one = select(meat, 
                  date, beef, veal, pork, lamb_and_mutton, broilers, other_chicken, turkey)
meat_melt = melt(as.data.frame(meat_one), id.vars = 'date')

#Time series graph

ggplot(meat_melt, aes(x = date, y = value, colour = variable)) +
  geom_line() +
  ggtitle('Trends over time')

#Lets smooth in order to see the trends:

ggplot(meat_melt, aes(x = date, y = value, colour = variable)) +
  geom_smooth() +
  ggtitle ('Trends over time, smoothed')

ggplot(meat_melt, aes(x = date, y = value)) + geom_smooth() +
  facet_wrap(~variable)